---
title: "MSBA Stats Project"
author: "Abigail Hoffman"
date: "2022-09-21"
output: html_document
---

```{r}
#Load Data 
df <- read.csv("Capital Bike Sharing data by hour.csv")

#Load Libraries 
library(tidyverse)
library(ggplot2)
library(magrittr)
library(broom)
library(corrplot)
library(stargazer)
#Check for Missing Values
apply(df,2,anyNA)

#Clean Data for Analysis
df <- mutate(df, month_name = (month.name[df$mnth]))
df$season <- as.numeric(df$season)


```
Fit a linear model predicting the total bike rental demand from daily temperature. What kind of insights can you generate? (make sure to write the linear model and interpret it in the context of the data) (20 points)

```{r}
est_bike <- lm(cnt~ temp, data=df)
tidy(est_bike)
glance(est_bike)

summary(est_bike)
```

Fit another linear model predicting total daily bike rentals from daily feeling temperature. Write the linear model, interpret the slope, etc. Is the temperature or feeling temperature a better predictor of bike rentals? (20 points)

```{r}
est_bike2 <- lm(cnt~ atemp, data=df)
tidy(est_bike2)
glance(est_bike2)

summary(est_bike2)



stargazer(est_bike, est_bike2, type="text",
          dep.var.caption = "", dep.var.labels.include = F,
          report = "vc*", df = F, model.numbers = F,single.row = T,
          keep.stat = c("ser","rsq","adj.rsq"),
          column.labels = c("Reg Temp","Reg Felt Temp"))
```


```{r}
#Convert Normalized Temperatures
tempconversion <- function(min,max,vector){
  result <- vector * (max-min)+ min
  return(result)
}

df$temp <- tempconversion(-8,39,df$temp)
df$atemp <- tempconversion(-16,50,df$atemp)
df$actual_windspeed <- df$windspeed*67
df$actual_humidity <- df$hum*100

#Edit variables for readability

#seasons
df$season <- factor(format(df$season, format="%A"),
                          levels = c("1", "2","3","4") , labels = c("Spring","Summer","Fall","Winter"))

#Working Day vs Holiday/Nonworking Day
df$holiday <- factor(format(df$holiday, format="%A"),
                          levels = c("0", "1") , labels = c("Working Day","Weekend or Holiday"))

#Weather type 
df$weathersit <- factor(format(df$weathersit, format="%A"),
                          levels = c("1", "2","3","4"), 
               labels = c("Good:Clear/Sunny","Moderate:Cloudy/Mist","Bad: Rain/Snow/Fog","Worse:Heavy Rain/Snow/Fog"))



```

No missing values in data set.

What is the trend in overall bike demand over the months of the year? (5 points)

```{r}
demand <- df %>% 
  group_by(mnth) %>% 
  summarise(bikedemand = sum(cnt))


#Line graph of total demand over 2011 and 2012 combined
plot(demand, type="l", 
      main = "Total Bike Rentals Over the Months of the Year",
     xlab = "Month",
     ylab = "Total Bike Rentals",
     col  = "steelblue2",
     pch  = 19)
  
```

The data science group at Capital bike share hypothesize that a. There must be high demand during office timings. Early morning and late evening can have different trends (cyclist) and low demand from 10:00pm to 4:00am. Do you agree?

```{r}
df %>% group_by(hr) %>%
  summarise(Hourly_Demand =mean(cnt)) %>%
  ggplot(aes(as.factor(hr), Hourly_Demand)) + 
  geom_line(group=1, col="steelblue2") +
  labs(title = "Hourly Average Demand of Bikes",
       subtitle= "Data from 2011-2012 in Washington, DC") +
  xlab(label="Hour of the Day (Military Time)") + 
  ylab(label= "Average Demand") + 
  theme_classic()




  
```

b.  Registered users demand more bike on weekdays compared to the weekend or holiday. Do you agree? (20 points)

```{r}

#total number
table(df$holiday)

#Percentages of each
mean(df$workingday)*100
mean(df$workingday==0)*100

#Totals comparison
df %>% 
  group_by(holiday) %>% 
  summarise(`Demand of Registered Users` = sum(registered),`Demand of Casual Users` = sum(casual),
            `Average Demand of Registered Users` = mean(registered),`Average Demand of Casual Users` = mean(casual))

#Take Average as there are more days of the week than weekend or holiday throughout the year

#casual comparison



```

I agree with the statement.

The table demonstrates that there is more demand on a working day.

Is there any relationship between season and bike rental? Create a visualization displaying the relationship. (10 points)

```{r}

table(df$season)

ggplot(df, aes(season, cnt, fill=season)) + 
  geom_boxplot(alpha=0.75) + 
   labs(x="",
       y="Demand",
       fill="",
       title="Bike Rental Demand By Season",
       subtitle="Data from 2011 to 2012 in Washington, DC") + 
   theme(plot.title = element_text(hjust=.5,size=15, color="steelblue4"),
        plot.subtitle = element_text(hjust=.5, size=15, color="steelblue2"))+
  scale_fill_manual(values=c("thistle2",
                             "snow1",
                             "azure3",
                             "slategray")) + 
  theme_classic()


```

What type of relationship do you see between weather and bike rental? Is the relationship the same for registered vs. casual users? (10 points)

```{r}

table(df$weathersit)



plot(df$temp, df$cnt ,type="p", col= 'pink', xlab = 'Actual Temperature', ylab = 'Total Bike Rentals')
ggplot(df, aes(x=temp,y=cnt))+geom_point(shape=1)+geom_smooth(method=lm)+ xlab("Actual Temp. in Celcius")+ylab("Bike Rentals")


Cor_actual_temp<-cor(x = df$temp, y = df$cnt)
Cor_actual_feel_temp <- cor(x = df$atemp, y =df$cnt)
df_cor<- df %>% select(cnt,temp,atemp,actual_humidity,actual_windspeed)
df_cor<- data.frame(df_cor)
  
colnames(df_cor)[1] <- "Total Rentals"
colnames(df_cor)[2] <- "Temperature"
colnames(df_cor)[3] <- "Feel Temp."
colnames(df_cor)[4] <- "Humidity"
colnames(df_cor)[5] <- "Windspeed"

cor(df_cor)
   

corplot_df <- cor(df_cor)
corrplot(corplot_df, method="number",tl.col = "black")

#registered Users

Cor_actual_temp<-cor(x = df$temp, y = df$registered)
Cor_actual_feel_temp <- cor(x = df$atemp, y =df$registered)
df_cor_r<- df %>% select(registered,temp,atemp,actual_humidity,actual_windspeed)
df_cor_r<- data.frame(df_cor_r)
  
colnames(df_cor_r)[1] <- "Total Reg. Rentals"
colnames(df_cor_r)[2] <- "Temperature"
colnames(df_cor_r)[3] <- "Feel Temp."
colnames(df_cor_r)[4] <- "Humidity"
colnames(df_cor_r)[5] <- "Windspeed"

cor(df_cor_r)
corplot_df_registered <- cor(df_cor_r)
corrplot(corplot_df_registered, method="number",tl.col = "black")
   
#casual users
Cor_actual_temp3<-cor(x = df$temp, y = df$casual)
Cor_actual_feel_temp3 <- cor(x = df$atemp, y =df$casual)
df_cor_casual<- df %>% select(casual,temp,atemp,actual_humidity,actual_windspeed)
df_cor_casual<- data.frame(df_cor_casual)
  
colnames(df_cor_casual)[1] <- "Total Casual Rentals"
colnames(df_cor_casual)[2] <- "Temperature"
colnames(df_cor_casual)[3] <- "Feel Temp."
colnames(df_cor_casual)[4] <- "Humidity"
colnames(df_cor_casual)[5] <- "Windspeed"

cor(df_cor_casual)
corplot_df_casual <- cor(df_cor_casual)
corrplot(corplot_df_casual, method="number",tl.col = "black")




   

```


```{r}
par(mfrow=c(1,3))
corrplot(corplot_df, method="number",tl.col = "black")
corrplot(corplot_df_registered, method="number",tl.col = "black")
corrplot(corplot_df_casual, method="number",tl.col = "black")

par(mfrow=c(1,1))
```



